How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-01 tse-paper/content.tex
explain techniques in intro, add running example and structural category example

@@

\dots

build logs valuable data source, but semi-structured and verbose,
therefore need (automated) processing

@@

To address this, we implement three promising techniques for build log analysis.
They retrieve specific chunks of targeted information and are configured by
examples provided by the user.
We conduct an empirical study on the \emph{LogChunks} \todo{cite} data set to
gauge the performance of these chunk retrieval techniques and characterize
when one of the techniques should be preferred over other alternatives (\textbf{RQ2}).

@@

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on 
  \item \textbf{a common text similarity approach (CTS)}
  \item \textbf{keyword search (KWS), an ad-hoc method for finding fitting passages by searching the whole text for the occurrence of a specific trigger word}
\end{itemize}
% TODO: Shortly describe each of these techniques

% TODO: Add running example?
\textbf{RQ2} asks under which conditions PBE, CTS and KWS are suited to retrieve information from continuous integration build logs.
\textbf{RQ2} is refined into sub-questions along with the criteria resulting from \textbf{RQ1} and compares their instantiations for the three techniques:
how many training examples a technique needs to perform best (\textbf{RQ2.1}), how structurally diverse the examples can be (\textbf{RQ2.2}) and how accurate the retrieved output is (\textbf{RQ2.3}).
To evaluate PBE, CTS and KWS we use the \emph{LogChunks} data set, which encompasses about 800 log files from 80 repositories.
% TODO: explain structural category based on running example from above?
Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural representation within the log.

@@

\noindent
\textbf{Our work contributes:}
\begin{itemize}
  \item A tool with three prototypical implementations of chunk retrieval techniques:
        \begin{itemize}
          \item program synthesis from examples using the Microsoft PROSE library (PBE),
          \item a common information retrieval approach using text similarity (CTS), and
          \item a keyword search approach (KWS).
        \end{itemize}
  \item Recommendations for the configuration of each of the investigated chunk retrieval techniques.
  \item Guidelines on choosing a suitable chunk retrieval technique.
\end{itemize}

@@

% TODO Caro cite guidelines
Since we cannot take for granted that ``analyzing build logs'' is the
cannonical term that other, isolated works have used, we instead
decdided to leave out the potentially variable ``analysis'' part and
instead focus on variations of the term ``build log''. This has the
advantage of being biunique: We explicitly decided to exclude systems
or event logs (see \Cref{sec:rw}). On the other hand, a Google Scholar
search for ``build log'' returns more than 2.2 million results
(2020-03-15). We therefore refined our search criterions for the four
different spelling possibilities of ``build log'' to exclude obvious
non-related works. This left us with 1,552 combined search results
(2020-03-15). The full search queries are documented in our replication
package \todo{cite/add replication package}.

@@

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
	basicstyle=\scriptsize,
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  keywordstyle=\color{blue},
	frame=single,
  extendedchars=false,
  texcl=false
}
\begin{figure*}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet- Responder: Received block blk_3587508140051953248 of size 67108864 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140051953248 terminating
  \end{lstlisting}  
  \caption{System Log excerpt. Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

@@

\lstset{
  language=caml,
	basicstyle=\scriptsize,
  morekeywords={PosToEndRegion, RegexPosition, RegexPair, StartToPosRegion},
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  stringstyle=\color{blue},
  keywordstyle=\bfseries\color{black},
	frame=single,
  extendedchars=false,
  escapeinside=//
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
let s = v in let s = PosToEndRegion(s, RegexPosition(s, RegexPair("Colon/$\circ$/Line Separator", "ALL CAPS"), 1)) in StartToPosRegion(s, RegexPosition(s, RegexPair("/$\varepsilon$/", "Line Separator/{\color{blue}$\circ$}/Line Separator"), 1))
  \end{lstlisting}  
  \caption{Text extraction program synthesized by FlashExtract.}
  \label{lst:prose-program}
\end{figure}

@@

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers scavenge for a specific piece of information within a large amount of unstructured information, a first ad-hoc approach they use is to search for related keywords.
Indeed, this was one of the most common approaches we took when searching for the reason the build failed within a log while creating the \emph{LogChunks} data set~\cite{logchunks}.
% As this is a technique readily available in many tools developers use to view build logs, we study when such a keyword search is suitable for retrieving information chunks from CI build logs.
% In the following we will explain how we use simple keyword search to retrieve information from CI build logs, which we refer to as \emph{KWS}\@.

@@

@@

\subsection{Other Techniques}
\label{sec:expl-rlr}
%Literature mentions further build log analysis techniques.
%This section describes them with our notion of chunk retrieval techniques.

@@

\lstset{
  morekeywords={Test Output, Desired Test Output},
	basicstyle=\scriptsize,
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  keywordstyle=\color{blue},
	frame=single,
  extendedchars=false,
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g. 1px, bold), was ";"
        on line 86 of sass/components/dropdown.sass   
Desired Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g. 1px, bold), was ";"
        on line 86 of sass/components/dropdown.sass
        from line 5 of sass/components/_all.sass
        from line 6 of bulma.sass
  \end{lstlisting}  
  \caption{Example for an unsuccessful retrieval (PBE retrieved only two of the four targeted lines).}
  \label{lst:pbe-unsuccessful}
\end{figure}

>>>>

% TODO It would be futile to compare our techniques with one of the existing manual regex parsers, as (a) as our literature survey showed, their scope is mainly limited to the Java ecosystem and (b) whether they would work or not would be randomly based on whether the project we select exhibits features their regular expressions were trained for. Finally, they are not really automatic

\lstset{
  morekeywords={},
	basicstyle=\ttfamily\scriptsize,
  postbreak=\mbox{\textcolor{blue}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  stringstyle=\color{Plum},
	frame=single,
  extendedchars=false,
  texcl=false,
  aboveskip=\baselineskip,
  belowskip=0pt
}

@@

@@

To address this, we implement three promising techniques for build log analysis.
They retrieve specific chunks of targeted information and are configured by
examples provided by the user.
We conduct an empirical study on the \emph{LogChunks} data set~\cite{brandt2020logchunks} to
gauge the performance of these chunk retrieval techniques and characterize
when one of the techniques should be preferred over other alternatives (\textbf{RQ2}).

@@

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on user examples of build logs and targeted log chunks, PBE synthesizes
  a regular expression matching the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  Using the Vector Space Model, CTS selects those lines of a log which are
  most similar to the lines present in example chunks provided by the user.
  \item \textbf{Keyword Search (KWS)}
  Ad-hoc method for finding fitting passages by searching the whole text for
  the occurrence of specific trigger words.
\end{itemize}

\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

* What went wrong:
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
Could not determine the dependencies of task `:app:jacocoTestDebugReport'.
> Task with path 'testDebug' not found in project ':app'.
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]

* Try:
  \end{lstlisting}  
  \caption{Example of a log chunk explaining why an Android/Gradle build failed.}
  % TODO Moritz: we can relate this to travistorrent maybe, you told me long ago that this was something your analyzer cannot extract
  \label{lst:chunk-example-1}
\end{figure}

\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

* What went wrong:
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
Execution failed for task ':app:connectedAndroidTest'.
> com.android.builder.testing.api.DeviceException: java.lang.RuntimeException: ...
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]

* Try:
  \end{lstlisting}  
  \caption{Another example of a log chunk explaining why an Android/Gradle build failed.}
  \label{lst:chunk-example-2}
\end{figure}

\lstset{
  language=caml,
  morekeywords={StartExtraction, RegexPosition, RegexPair, EndExtraction},
  keywordstyle=\bfseries\color{black},
  escapeinside=//
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
StartExtraction(RegexPair("Colon/{\color{Plum}$\circ$}/Line Separator", "ALL CAPS")) in EndExtraction(RegexPair("/{\color{Plum}$\varepsilon$}/", "Line Separator/{\color{Plum}$\circ$}/Line Separator"))
  \end{lstlisting}  
  \caption{Simiplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example-1,lst:chunk-example-2}}
  \label{lst:prose-program-simplified}
\end{figure}

\lstset{
  language=,
  morekeywords={},
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
--- FAIL: TestSeparator (0.03s)
    repo\_test.go:75: expected entry to be in form of `* [link] - description.`, got '* ...
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}  
  \caption{Example of a log chunk. \todo{which build tool?}
  \todo{move to where structural category explanation}}
  \label{lst:chunk-example-3}
\end{figure}

The results of our empirical comparison study on the chunk retrieval techniques
show that the structural representation of the targeted chunk greatly influences
the performance of the chunk retrieval techinques.
\Cref{lst:chunk-example-1,lst:chunk-example-2} show two exemplary log chunks
from Android/Gradle build logs. Both chunks contain the information on why the
corresponding build failed. When PBE is trained on these two examples it produces
a regular expression similar to \Cref{lst:prose-program-simplified}. The retrieval
starts after a colon followed by a line separator and before a capital letter.
It ends before two line separators, consistent with the two example chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2}.
This shows, that the characters structuring the text around the chunk are cruicial
for regular expressions to identify the targeted log chunk. In fact, if the third
example from \Cref{lst:chunk-example-3} with a different structural representation is added,
PBE is not able to syntesize a program and returns ``\texttt{no program found}''

% \textbf{RQ2} asks under which conditions PBE, CTS and KWS are suited to retrieve information from continuous integration build logs.
% \textbf{RQ2} is refined into sub-questions along with the criteria resulting from \textbf{RQ1} and compares their instantiations for the three techniques:
% how many training examples a technique needs to perform best (\textbf{RQ2.1}), how structurally diverse the examples can be (\textbf{RQ2.2}) and how accurate the retrieved output is (\textbf{RQ2.3}).
% To evaluate PBE, CTS and KWS we use the \emph{LogChunks} data set, which encompasses about 800 log files from 80 repositories.
% Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural representation within the log.

@@

% \noindent
% \textbf{Our work contributes:}
% \begin{itemize}
%   \item A tool with three prototypical implementations of chunk retrieval techniques:
%         \begin{itemize}
%           \item program synthesis from examples using the Microsoft PROSE library (PBE),
%           \item a common information retrieval approach using text similarity (CTS), and
%           \item a keyword search approach (KWS).
%         \end{itemize}
%   \item Recommendations for the configuration of each of the investigated chunk retrieval techniques.
%   \item Guidelines on choosing a suitable chunk retrieval technique.
% \end{itemize}

@@

% TODO Caro cite guidelines
Since we cannot take for granted that ``analyzing build logs'' is the
cannonical term that other, isolated works have used, we instead
decdided to leave out the potentially variable ``analysis'' part and
instead focus on variations of the term ``build log''. This has the
advantage of being biunique: We explicitly decided to exclude systems
or event logs (see \Cref{sec:rw}). On the other hand, a Google Scholar
search for ``build log'' returns more than 2.2 million results
(2020-03-15). We therefore refined our search criterions for the four
different spelling possibilities of ``build log'' to exclude obvious
non-related works. This left us with 1,552 combined search results
(2020-03-15). The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.

@@

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
  keywordstyle=\color{Plum},
  escapeinside=**
}
\begin{figure*}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet Responder: Received block blk_3587508140 of size 67108 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140 terminating
  \end{lstlisting}  
  \caption{System Log excerpt. Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

@@

\lstset{
  language=caml,
  morekeywords={PosToEndRegion, RegexPosition, RegexPair, StartToPosRegion},
  keywordstyle=\bfseries\color{black},
  escapeinside=//
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
let s = v in let s = PosToEndRegion(s, RegexPosition(s, RegexPair("Colon/$\circ$/Line Separator", "ALL CAPS"), 1)) in StartToPosRegion(s, RegexPosition(s, RegexPair("/$\varepsilon$/", "Line Separator/{\color{blue}$\circ$}/Line Separator"), 1))
  \end{lstlisting}  
  \caption{Text extraction program synthesized by FlashExtract.}
  \label{lst:prose-program}
\end{figure}

@@

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers scavenge for a specific piece of information within a large amount of unstructured information, a first ad-hoc approach they use is to search for related keywords.
Indeed, this was one of the most common approaches we took when searching for the reason the build failed within a log while creating the \emph{LogChunks} data set~\cite{brandt2020logchunks}.
% As this is a technique readily available in many tools developers use to view build logs, we study when such a keyword search is suitable for retrieving information chunks from CI build logs.
% In the following we will explain how we use simple keyword search to retrieve information from CI build logs, which we refer to as \emph{KWS}\@.

@@

\subsection{Random Line Retrieval}
\label{sec:expl-rlr}

@@

\subsection{Other Techniques}
%Literature mentions further build log analysis techniques.
%This section describes them with our notion of chunk retrieval techniques.

@@

\lstset{
  language=,
  morekeywords={Test, Output, Desired},
	basicstyle=\ttfamily\scriptsize,
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  keywordstyle=\textbf,
	frame=single,
  extendedchars=false,
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g. 1px, bold), was ";"
        on line 86 of sass/components/dropdown.sass   
Desired Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g. 1px, bold), was ";"
        on line 86 of sass/components/dropdown.sass
        from line 5 of sass/components/_all.sass
        from line 6 of bulma.sass
  \end{lstlisting}  
  \caption{Example for an unsuccessful retrieval (PBE retrieved only two of the four targeted lines).}
  \label{lst:pbe-unsuccessful}
\end{figure}
